Scaling laws of trophic level and generality vary across latitude

### 2014-06-19 manuscript/latitudepaper.tex
Aggregated lakes and streams as 'freshwaters' for simpler analysis.

We then converted each web to a trophic-species version by aggregating species
with the same predators and prey. This helps to reduce variation in resolution
across different studies.
We then classified them both according to latitude
and ecosystem type (terrestrial, stream, lake, marine, estuary). We then calculated 17
food-web properties for each web (Table 1).

@@

For the food-web properties that did scale with species-richness or connectance, most of these relationships were 
constant over latitude. Latitude terms were not included in the best-fit models for vulnerability, SD of generality or 
SD of links. SD of vulnerability showed a weak negative relationship with latitude ($\beta$=-0.001, $p$=0.056), 
and path length increased with increasing 
latitude ($\beta$ = 0.005, $p$=0.045) but neither latitude effect affected any scaling with species richness or 
connectance.

The scaling rules of mean and maximum trophic level and generality varied over latitude. 
Maximum trophic level increased with 
latitude ($\beta=0.011$, $p$=0.011), but this relationship weakened as species richness increased 
($\beta$=-0.008, $p$=0.011). Similarly, maximum trophic level increased with species richness ($\beta$=0.430, $p$=0.003) 
except at very high latitudes. Maximum trophic level decreased with connectance for estuarine food webs 
($\beta$=-0.458, $p$=0.021), varied little in lakes ($\beta$=-0.127, $p$=0.192), and increased with connectance in
marine, stream, and terrestrial 
food webs ($\beta$=0.219, $p$=0.005; $\beta$=0.249, $p$<0.001; and $\beta$=0.337, $p$=0.002, respectively). Taken 
together, these effects resulted in an increase in maximum trophic level with latitude at low species richnesses across 
all ecotypes and decreasing maximum trophic level with latitude at high species richness across all ecotypes 
(Fig. \ref{}). At moderate levels of species richness, trophic level increased with latitude for estuarine webs, showed 
little change for lake webs, and decreased for all other ecotypes.

Qualitatively similar trends were observed for mean trophic level. Mean trophic level increased with latitude 
($\beta$=0.0109, $p$<0.001) and with species richness ($\beta$=0.616, $p$<0.001), but both trends were weakened by an 
interaction between latitude and species-richness ($\beta$=-0.008, $p$<0.001). The best-fitting model for mean trophic 
level did not include a connectance term. The increase in mean trophic level with species richness was highest for 
estuarine food webs ($\beta$=0.616, $p$<0.001) and lower for lakes ($\beta$=-0.608, $p$=0.232), streams 
($\beta$=0.238, $p$<0.001), terrestrial food webs ($\beta$=0.217, $p$=0.002), and marine food webs 
($\beta$=0.133, $p$<0.001).

Generality showed the opposite trend, decreasing with increasing latitude ($\beta$=-0.017, $p$=0.003). This decrease 
was counteracted by a positive interaction between latitude and species richness ($\beta$=0.013, $p$=0.003), with the net 
result that generality increased with latitude for most ecotypes and species richnesses (Fig. \ref{}).
The main effect of species richness on generality varied across ecotypes. Generality decreased with increasing species 
richness for estuarine webs ($\beta$=-0.058, $p$=0.003) and increased for lakes ($\beta$=0.220, $p$=0.245), 
streams ($\beta$=0.543, $p$<0.001), terrestrial ($\beta$=0.603, $p$=0.005) and marine webs ($\beta$=0.761, $p$<0.001). 
Generality increased with connectance equally for all ecotypes ($\beta$=0.875, $p$<0.001).

@@

The lack of latitudinal gradients in most of the food-web properties considered here may be due to the lack of a latitudinal 
gradient in species richness in this data set. If so, that strongly implies that the latitudinal gradients that were observed
were driven by some other factor than latitudinal changes in species richness. As we observed increasing generality away from 
the equator in conjunction with increasing trophic level, it appears that high-trophic level species are also more apt to be 
generalists. Generalism may be an important strategy for high-level consumers as their prey may be scarce or 
seasonal \citep{}. These results also 
suggest that the tropics may indeed have narrower niches \citep{}. Similar increases in geographic range size with 
latitude (Rapoport's rule) have been found for mammals \citep{Letcher1994}, XX \citep{}, and XX \citep{}. Large 
geographic ranges imply climactic generalization \citep{Letcher1994}, but may also promote dietary generalization 
(i.e., a species with a wide range may be more likely to eat a variety of different prey depending on what is 
available). This is supported by the fact tha feeding on many prey within a trophic level (non-omnivorous generality) 
has been shown to decrease population variability in zooplankton \citep{Romanuk2006}.

This contrasts with an earlier study of plant-animal mutualistic networks which found that generality increased in the tropics
\citep{Schleuning2012}. That increase in generality was attributed to increasing species richness, as high resource diversity 
might lead to generalized consumers \citep{Schleuning2012}. Unfortunately, this study did not investigate the scaling 
relationship of species richness with generality, so it is not clear whether a latitudinal gradient in species richness was 
responsible for the latitudinal gradient in generality. Regardless, the opposite trends in generality between antagonistic 
and mutualistic networks may indicate that different types of interactions between species are under different types of pressures
and hence respond differently to changes in temperature, productivity, or other latitudinally-associated variables.

@@

\begin{center}
\begin{table}[!h]
\caption{Symbols for and descriptions of food-web properties.}
\label{symbols}
\begin{tabular}{c l l}
\hline
Symbol      & Name & Description \\
\hline
S & Species richness & Number of trophic species \\
LS & Links per species & Mean number of links per taxon \\
C & Connectance & Number of observed links divided by the number of possible links ($L/S^{2}$) \\
\%Top & \% Top predators & Percentage of taxa without consumers \\
\%Int & \% Intermediate consumers & Percentage of taxa with both consumers and resources \\
\%Basal & \% Basal Resources & Percentage of taxa without resources \\
\%Herb	& \% Herbivores	& Percentage of herbivores and detrivores (taxa that feed on basal taxa) \\
\%Omni & \%Omnivores & Percentage of omnivores (taxa that feed on ≥2 taxa with different trophic levels) \\
Gen & Generality & Mean number of resources per taxon \\
GenSD & SD of generality & Standard deviation of resources per taxon \\
Vul & Vulnerability & Mean number of consumers per taxon \\
VulSD & SD of vulnerability & Standard deviation of consumers per taxon \\
LinkSD & SD of links per species & Standard deviation of links (consumers and resources) per taxon \\
SWTL & Mean short-weighted trophic level & Short-weighted trophic level averaged across taxa \\
Path & Path length & Characteristic path length, the mean shortest food chain between species pairs \\
\hline
\end{tabular}
\end{table}
% NODF & Average Nestedness & Degree to which interactions among specialists are subsets or interactions among generalists \\
% Chain & Chain length & Mean food chain length \\
% Cluster & Clustering coefficient & Mean clustering coefficient (probability that two taxa linked to the same taxon are also linked) \\

>>>>

We then converted each web to a trophic-species version by aggregating species
with the same predators and prey. This helps to reduce variation in resolution
across different studies.
We then classified them both according to latitude
and ecosystem type (terrestrial, freshwater, marine, or estuary). We then calculated 17
food-web properties for each web (Table \ref{symboltable}).

@@

For the food-web properties that did scale with species-richness or connectance, most of these relationships were 
constant over latitude. Latitude terms were not included in the best-fit models for SD of generality or 
SD of links. Best-fit models for vulnerability, SD of vulnerability, and path length included latitude terms but showed 
very weak relationships ($\beta$=3.52 x $10^{-17}$, $p$=0.170; $\beta$=-0.001, $p$=0.050; and $\beta$=0.005, $p$=0.069; 
respectively). None of the above models included an effect of latitude on scaling with S or C.

The scaling rules of mean trophic level and generality, however, did vary over latitude. 
Mean trophic level increased with latitude ($\beta$=0.011, $p$<0.001) and with species richness 
($\beta$=0.633, $p$<0.001), but both trends were weakened by an interaction between latitude and species-richness 
($\beta$=-0.008, $p$=0.001). The best-fitting model for mean trophic 
level did not include a connectance term. The increase in mean trophic level with species richness was highest for 
estuarine food webs ($\beta$=0.633, $p$<0.001) and lower for freshwater ($\beta$=0.259, $p$=0.002), terrestrial 
($\beta$=0.234, $p$=0.005), and marine food webs ($\beta$=0.152, $p$<0.001). For food webs with high species richness, 
trophic level declined with latitude while trophic level in species-poor webs showed the opposite trend 
(Fig. \ref{TLfig}). The number of species for which the slope of trophic level over latitude changed from negative to 
positive varied across ecosystem types.

Generality showed the opposite trend, increasing with latitude for webs with high species richness and decreasing with
latitude for species-poor webs (Fig. \ref{genfig}).
This increase was driven by a positive interaction between latitude and species richness ($\beta$=0.014, $p$=0.002), as 
the main effect of latitude was negative ($\beta$=-0.018, $p$=0.002). Generality also decreased with species richness 
for estuarine food webs ($\beta$=-0.150, $p$=0.002) but increased with species richness for freshwater, terrestrial, 
and marine webs ($\beta$=0.453, $p$=0.005; $\beta$=0.523, $p$=0.006; and $\beta$=0.675, $p$=0.001; respectively). 
Generality increased with connectance equally in all ecotypes and across all latitudes ($\beta$=0.799, $p$<0.001).

@@

The lack of latitudinal gradients in most of the food-web properties considered here may be due to the lack of a 
latitudinal gradient in species richness in this data set. If so, that strongly implies that the latitudinal gradients 
that were observed were driven by some other factor than latitudinal changes in species richness. As we observed 
increasing generality away from 
the equator in conjunction with decreasing trophic level (at least in high-S food webs), it appears that species nearer
the top of the food web may be feeding on a wider variety of prey at the poles, lowering their trophic levels. 
As both feeding on many prey within a trophic level (non-omnivorous generality) \citep{Romanuk2006} and feeding on many 
trophic levels \citep{} have been shown to decrease population variability, high generality at higher latitudes may act 
as ``insurance'' for upper-level consumers by allowing them to switch prey according to availability.

This contrasts with an earlier study of plant-animal mutualistic networks which found that generality increased in the 
tropics \citep{Schleuning2012}. That increase in generality was attributed to increasing species richness, as high 
resource diversity might lead to generalized consumers \citep{Schleuning2012}. While we also found increasing 
generality with increasing species richness in our dataset, it seems unlikely that higher species richness at the poles 
is driving the observed trend in generality. This suggests that while mutualistic and antagonistic interactions may
be similarly affected by species richness, they respond differently to other latitudinally-associated variables.

[[High TL/low generality in species-poor food webs?]]

@@

\begin{center}
\begin{table}[!h]
\caption{Symbols for and descriptions of food-web properties.}
\label{symboltable}
\begin{tabular}{c l l}
\hline
Symbol      & Name & Description \\
\hline
S & Species richness & Number of trophic species \\
LS & Links per species & Mean number of links per taxon \\
C & Connectance & Number of observed links divided by the number of possible links ($L/S^{2}$) \\
\%Top & \% Top predators & Percentage of taxa without consumers \\
\%Int & \% Intermediate consumers & Percentage of taxa with both consumers and resources \\
\%Basal & \% Basal Resources & Percentage of taxa without resources \\
\%Herb	& \% Herbivores	& Percentage of herbivores and detrivores (taxa that feed on basal taxa) \\
\%Omni & \%Omnivores & Percentage of omnivores (taxa that feed on ≥2 taxa with different trophic levels) \\
Gen & Generality & Mean number of resources per taxon \\
GenSD & SD of generality & Standard deviation of resources per taxon \\
Vul & Vulnerability & Mean number of consumers per taxon \\
VulSD & SD of vulnerability & Standard deviation of consumers per taxon \\
LinkSD & SD of links per species & Standard deviation of links (consumers and resources) per taxon \\
SWTL & Mean short-weighted trophic level & Short-weighted trophic level averaged across taxa \\
Path & Path length & Characteristic path length, the mean shortest food chain between species pairs \\
\hline
\end{tabular}
\end{table}
% NODF & Average Nestedness & Degree to which interactions among specialists are subsets or interactions among generalists \\
% Chain & Chain length & Mean food chain length \\
% Cluster & Clustering coefficient & Mean clustering coefficient (probability that two taxa linked to the same taxon are also linked) \\
